A graph-based data model for interactive evolutionary computation

### 2017-01-19 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
About to start explaining the model

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
This paper presents an alternative data model for the development of
Interactive Evolutionary Computation (IEC) algorithms. A graph %specify
                                %the elements of the graph - JJ
stored for efficiency in a graph database, is
proposed to model the overall system which consists of: a social network of
volunteers participating in the interactive evolutionary experiment, a
population of individuals in the evolutionary algorithm, 
the system\'s functionality and their interactions. This graph model can then be used
to tackle some of the issues associated with IEC such as, 
user fatigue, lack of volunteer participation, fitness estimation % what is this issue?
%Maybe the abstract should start with the issues we want to tackle as
%motivation. Motivation always goes first - JJ
and novelty. % These two last issues are not too clear, maybe they
             % could be explained in that motivation. 
The proposed graph database could be used to model IEC systems that 
follow a user-centered approach. % Graph is the model, graph database
                                % the implementation. Two different
                                % things. 
A case study is presented providing both 
conceptual and implementation details using the EvoSpace-Interactive
platform.
% It must prove that the issues have been addressed - JJ
\end{abstract}

@@

\section{Introduction}
% Maybe an introduction on why and where to use these types of
% algorithms? - jj
% Done -Mario
Interactive Evolutionary Algorithms (IEAs) are standard EAs whose
fitness evaluations are performed by persons within an interactive 
system.  Thus, the main loop of the EA needs the intervention of a
human for the quality assessment of each proposed solution.
IEAs have demonstrated their ability for effectively
producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992},
as well as other products in many other domains \cite{ie1}. 
Several IECs have been developed as web based systems \cite{picbreeder},
who depend on volunteers who visit the system to interactively help
with the search, using both anonymous and registered users.
These systems have several inherent drawbacks arising from the very nature of 
the algorithms, namely, the human fatigue caused by the interaction, and
the boredom arising when users evaluate a large number of individuals 
many of which are not interesting or are very similar to each other.
IEAs also share some of the issues commonly found in volunteer based systems
\cite{sarmenta2001volunteer,web:BOINC} such as the volunteer\'s lack of accountability,
and the need to build trust between participants and application providers. 
For the provider there is also the difficulty of establishing 
the amount of time and resources
a volunteer is willing to spend on the system, and how they decide if they
participate or not \cite{JJ:2016}. In order to increase participation and 
engagement, and to tackle the issues mentioned above,  
IEAs must follow a user centered design, giving extensive attention 
to volunteer users, not only because their
explicit evaluation is essential, but also because the context of the 
interaction affects the system as a whole. Users interact not only with the graphical
interface, and each individual\'s phenotype, there is also the interaction
with other users, their previous evaluations and their own stream of activities
and past experiences. Thus users in this proposal are defined as dynamic entities 
interacting with an evolutionary computation having the fundamental purpose 
of evaluating individuals according to their preferences,
both explicitly or implicitly. In this work we propose giving the same
importance to users and their interactions as the population of 
individuals have in a traditional EA, even to the extend as using this
knowledge as an integral part of the algorithm it self. We argue also
that a graph model is a viable representation for this purpose,  
presenting a natural way of maping the relationships found in the system. 
Most IEAs can have the following many-to-many relations:

\begin{itemize}
  \item { \tt Individual-Individual} The relationship between individuals 
  is naturally expressed as a tree, or a graph when both ancestors 
  and descendants is considered. Other semantic relations can also be considered
  for instance: similarity or composition.

  \item {\tt User-Individual}
  Many kinds of relationships between users and individuals,
  can be expressed using common actions found in social networks;
  for instance share, like, rate or add to collection [cite Activity Streams 2.0
  syntax (https://www.w3.org/TR/activitystreams-vocabulary)]. In IEAs these
  relationships are used to assigned a single or a collection of
  fitness values to individuals [cite ES-I].

  \item {\tt User-User}
  Relationship between users can also be modeled 
  as a social network, with well stablished algorithms and metrics (cite).
  Identifyng leaders of opinion and the similarity between user\'s 
  preferences is of special importance for IEAs, as they could be used
  to recommend to users only those individuals they will probably like more.

@@

A case study is presented as a proof of concept, providing both 
conceptual and implementation details of a graph data model for
the implementation of an IEA in the EvoSpace-Interactive platform. 
The graph database is then used to measure the degree of relationship
between users and their participation.

@@

@@

\begin{figure*}
\captionsetup{justification=centering,margin=2cm}
\centering
\setlength\fboxsep{0pt}
\setlength\fboxrule{0.7pt}
\fbox{\includegraphics[width=12cm,height=10cm,keepaspectratio]{img/UI_ed01.png}}
\caption{User interface ED01.}
\label{fig:web}       
\end{figure*}

@@

\section{IEC Graph based model}
\label{sec:graph}
The fundamental goal of this research is to develop a user-centered framework
for interactive evolutionary computation (IEC) in order to increase users’
participation and also to minimize the amount of evaluations needed for the
evolutionary process in given Web-based IEC application.

In this chapter an explanation of the proposed framework is described. The
different techniques used, such as user modeling, fuzzy logic, and human-
interaction.

This framework is presented in figure \ref{fig:uc_framework}. Each of the
entities of this framework will be explained in detail in following sections.

The current model is tailored to the EvoSpace-I functionality and capabilities
but it can be modified to handle other platforms, just by adding new clases 
of nodes.

@@

@@

\subsection{Users}

@@

In order to start the task of evaluating individuals is necessary that users
access the system through a login mechanism. This mechanism consists of
providing a username and a password in order to grant access to the system. All
users accessing in this way they are considered active users within an IEC
system.

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
Interactive evolution is the leading technique employed by 
Evolutionary Art researchers, where users' preferences guide the search. A common
practice is to develop Interactive Evolutionary Computation (IEC) algorithms
as web applications in order to give users easy access the application as they
often depend on volunteers who visit the system to interactively help with the search. 
There are several drawbacks that limit user participation 
such as human fatigue and boredom resulting from evaluating 
a large number of individuals. 
To tackle this issues, we propose the design of IEC systems using a user-centered 
approach, with a framework consisting of a social network of 
volunteers, interacting with a population consisting again of a network 
of individuals. In this work we propose a graph model as a viable 
and efficient tool for mapping in a natural way the relationships 
found in the system. A case study is presented as a proof of concept, 
providing both conceptual and implementation details of a graph based model as it is 
applied in the implementation of an IEA using the EvoSpace-Interactive platform. 
As a result, the graph model was then succesfully applied to measure the degree
of relationship between users with the porpuse of quantifing 
their participation.

% what is this issue?
%Maybe the abstract should start with the issues we want to tackle as
%motivation. Motivation always goes first - JJ
%and novelty. % These two last issues are not too clear, maybe they
             % could be explained in that motivation. 
%The proposed graph database could be used to model IEC systems that 
%follow a user-centered approach. % Graph is the model, graph database
                                % the implementation. Two different
                                % things.

% It must prove that the issues have been addressed - JJ
\end{abstract}

@@

\section{Introduction}
% Maybe an introduction on why and where to use these types of
% algorithms? - jj
% Done -Mario
Interactive Evolutionary Algorithms (IEAs) are standard EAs whose
fitness evaluations are performed by persons within an interactive 
system.  Thus, the main loop of the EA needs the intervention of a
human for the quality assessment of each proposed solution.
IEAs have demonstrated their ability for effectively
producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992},
as well as other products in many other domains \cite{ie1}. 
Several IECs have been developed as web based systems \cite{picbreeder},
who depend on volunteers who visit the system to interactively help
with the search, using both anonymous and registered users.
These systems have several inherent drawbacks arising from the very nature of 
the algorithms, namely, the human fatigue caused by the interaction, and
the boredom arising when users evaluate a large number of individuals 
many of which are not interesting or are very similar to each other.
IEAs also share some of the issues commonly found in volunteer based systems
\cite{sarmenta2001volunteer,web:BOINC} such as the volunteer\'s lack of accountability,
and the need to build trust between participants and application providers. 
For the provider there is also the difficulty of establishing 
the amount of time and resources
a volunteer is willing to spend on the system, and how they decide if they
participate or not \cite{JJ:2016}. In order to increase participation and 
engagement, and to tackle the issues mentioned above,  
IEAs must follow a user centered design, giving extensive attention 
to volunteer users, not only because their
explicit evaluation is essential, but also because the context of the 
interaction affects the system as a whole. Users interact not only with the graphical
interface, and each individual\'s phenotype, there is also the interaction
with other users, their previous evaluations and their own stream of activities
and past experiences. Thus users in this proposal are defined as dynamic entities 
interacting with an evolutionary computation having the fundamental purpose 
of evaluating individuals according to their preferences,
both explicitly or implicitly. In this work we propose giving the same
importance to users and their interactions as the population of 
individuals have in a traditional EA, even to the extend as using this
knowledge as an integral part of the evolutionary algorithm. We argue also
that a graph model is a viable representation for this purpose,  
presenting a natural way of maping the relationships found in the system. 
Most IEAs have the following many-to-many relations:

\begin{itemize}
  \item { \tt Individual-Individual} The relationships between individuals in the
  population are naturally expressed as a tree, or a graph when both ancestors 
  and descendants are considered. Other semantic relations can also be of interest
  for instance similarity or composition.

  \item {\tt User-Individual}
  Many kind of relationships between users and individuals,
  can be expressed using common actions found in social networks;
  for instance: share, like, rate or add to collection [cite Activity Streams 2.0
  syntax (https://www.w3.org/TR/activitystreams-vocabulary)]. In IEAs these
  relationships are used to assign a single or a collection of
  fitness values to individuals [cite ES-I].

  \item {\tt User-User}
  Relationship between users can also be modeled 
  as a social network, with well stablished semantics, algorithms and metrics (cite).
  A graph model could eneable researchers to find other ways of identifyng leaders of 
  opinion or measuring the similarity between user\'s preferences. 
  These measures can then be used by recommender algorithms selecting 
  individuals according users\' preferences.

@@

A case study is presented as a proof of concept, providing both
conceptual and implementation details of a graph data model applied in the 
implementation of an IEA. The experiment was implmented using
the EvoSpace-Interactive framework, and consisted of a web application
for evolving artistic drawings. The graph database is then used to
measure the degree of relationship between users and their participation.

@@

\begin{figure*}
\captionsetup{justification=centering,margin=2cm}
\centering
\setlength\fboxsep{0pt}
\setlength\fboxrule{0.7pt}
\fbox{\includegraphics[width=12cm,height=10cm,keepaspectratio]{img/UI_ed01.png}}
\caption{User interface of an EvoSpace-Interactive application.}
\label{fig:web}       
\end{figure*}

@@

@@

\section{IEC Graph based model}
\label{sec:graph}
The fundamental goal of this research is to develop a user-centered framework
for interactive evolutionary computation (IEC) in order to increase users’
participation and also to minimize the amount of evaluations needed for the
evolutionary process in given Web-based IEC application.  This framework is 
presented in figure \ref{fig:uc_framework}. The current model is tailored 
to the EvoSpace-I functionality and capabilities but it can be modified to 
handle other platforms, just by adding new clases of nodes.

@@

The user centered framework consists of three high level components:

\begin{itemize}
  \item { \tt Interactive Evolutionary Computational system.}  
  This is the real world system that we are going to model, it consist of
  human users and their interaction with one or more individuals from
  the population, this could be through a device, user interface or
  by directly interacting with real world objects generated as part of the algorithm.

  \item {\tt IEC Graph}
  This is the model proposed in this work: A graph database modeling
  the IEC system. The  main components are the social network of users and the population. This database is the back-end of the system, is used by the EA and
  the user interface.

  \item {\tt Evolutionary Algorithm} 
  This algorithm queries the graph model, and then updates those nodes that 
  are part of the population. 
\end{itemize}

@@

\subsection{IEC Graph Model}

In this section the Graph Model is presented,

@@
